Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds

### 2015-12-17 paper.tex
about to clean up

\title
{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds
%Unifying quasars with clumpy wind models
}

@@

Monte Carlo radiative transfer and photoionization code PYTHON.  Previously, we have shown that we could produce synthetic spectra resembling those of observed AGN for plausible wind models, but only if the X-ray luminosity was limited to about  10$^{43}$ erg s$^{-1}$.  Here , we investigate the  degree to which  clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame UV that have the characteristics of quasars with X-ray luminosities as large as 10$^{45}$ erg s$^{-1}$. 
We conclude, perhaps not surprisingly,  that clumping does maintain the ionization state of the wind necessary for strong BAL features in the rest-frame UV at more realistic X-ray luminosities. We examine the X-ray properties of these simple clumped models

@@

excited line emission in, e.g., \civ\ and \la, to emerge at the low inclination, 
`Type 1 quasar-like' angles. At the highest inclinations, the synthetic spectra possess prominent \mg\ and \al\ BALs, the absorption features seen in LoBAL quasars. 
Despite these successes, we are unable to reproduce 
the remarkably uniform emission line properties 
seen in BAL and non-BAL quasar composites, as we find a marked increase
in emission line equivalent widths at high inclination `BALQSO-like' angles. 
This is due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. 

@@

ionization states seen in AGN.
This suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a specific challenge to models in which an equatorial outflow
rises from a limb-darkened, foreshortened accretion disc.

@@

Due to their ubiquitous nature,
disc winds offer a natural explanation for the
diverse phenemonology of luminous AGN and quasars \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this {\em geometric unification} framework, the broad-line region (BLR) can 
correspond either to the dense wind base or clumps embedded
in the outflow. Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

@@

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BALQSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer (MCRT) code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13]{simlong2008,sim2010,higginbottom2013}.  The results have been encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BALQSOs, as long as the luminosity of the X-ray source was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher X-ray luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

@@

For this study, we use the MCRT code \py 
\footnote{Named {\sl c. 1995}, predating the inexorable rise of a certain progamming language.} we have developed to carry out our 
radiative transfer and photoionization simulations in non-local-thermodynamic-equilibrium 
(non-LTE). The code can be used to model a variety of
disc-wind systems; it has been used with application to accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The code operates as follows:   This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry. From some initial conditions in each cell, the code first calculates the ionization structure of the wind in a series of iterations.  Each iteration in an ``ionization cycle'' consists of generating photons, actually photon packets, from an accretion disc
and central object, and calculating how these photon bundles scatter through the wind (eventually escaping the outflow or hitting the disk). Then updating the ionization structure based on the properties of the radiation field in each cell, and the process is repeated.   Once the ionization structure has converged, the ionization structure is held fixed, and  synthetic spectra are generated at specific inclination angles in a series of "spectral cycles". LK02 provide a more detailed  description of the original code; various improvements have been made since then and are described by \cite{simmacro2005}, H13 and M15.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made for this study intended to improve the ionization calculation of H and He
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}). 
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15.
%(an improvement over that used by H14 for our earlier quasar study), 
In this scheme, the energy packets interact with either two-level `simple ions' 
or full `macro-atoms'. This allows one to treat non-LTE line transfer in radiative equilibrium, which assumes both statistical equilibrium and that 
radiative heating balances radiative cooling.
% {\bf KSL: what does "in radiative equilibrium" here mean? 
% JM: Defined.}
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple-ions (see M15). In this study,
only H and He are treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, 
and because we are especially interested in the contribution to 
AGN spectra of \LA.  
H13 treated all atoms in a two-level approximation.  
% {\bf KSL:
% I thought we were also doing He. JM: yep, corrected.}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
This allows the calculation of a photoionization rate estimator. Ion abundances are
then calculated by solving the rate equations between ions. We include collisional ionization and photoionization balanced with radiative, 
dielectronic and collisional (three-body) recombination.
As in M15, we use a dilute Boltzmann approximation to calculate 
the populations of levels for simple-ions. We stress that this approximation 
is not required for macro-atom levels. 
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
% The ionization rate out of ion $j$ can then be written as 
% \begin{equation}
% R_{j,j+1}(J)= 
% \displaystyle{
% n_j \left(C_{j} n_e + 
% \sum_{band~i=0}^{n}~
% {\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
% \right),}
% \end{equation}
% where $\sigma_j$ is the photoionization cross-section, $n_e$
% is the electron density and $C_{j}$
% represents the collisional ionization coefficient.
% The recombination rate {\em into} ion $j$ is given by 
% \begin{equation}
% R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
% \end{equation}
% where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
% The subscripts denote radiative and 
% collisional (three-body) recombination, respectively. 
% $T_e$ is the electron temperature.
% %% We neglect recombination to and from excited states in the simple-ion calculation. 
% For simple-ions, we use a dilute Boltzmann approximation to calculate 
% the population of level $k$ in ionic stage $j$,
% \begin{equation}
% \frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
% \end{equation}
% Here $z_j$ is the partition function of ionic stage $j$,
% $T_R$ is the effective radiation temperature,
% $E_k$ is the energy difference between level $k$ and the ground state,
% and $g_k$ is the statistical weight of level $k$. 
% We stress that this approximation is not required for ions
% treated as macro-atoms. 
% {\bf KSL: 
% It seems a bit odd that you explain what we do with simple ions our old approach, but don't say what is done for macro-atoms,.  You only say it is not the other.
% JM: We've shortened the section on simple-ions as agreed..}

@@

In our previous modelling efforts, we have assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
A clumpy outflow offers a possible solution to the so-called `over-ionization problem' in 
quasar and AGN outflows \citep{hamann2013}. This is the main motivation for incorporating 
clumping into our model.

Implementing a treatment of clumping is challenging, for
two main reasons. First, the physical scale lengths and density contrasts 
associated with these parameters are not well-constrained from observations.  
Second, there are significant computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. 
% Second, the addition of multiple additional degrees of
% freedom in the model results in significantly wider parameter space.
% Unfortunately, the physical scale lengths and density contrasts 
% associated with these parameters are not well-constrained from observations.  
% {\bf KSL: Reverse the order of these difficulties.  The most important is that it is not clear what kind of clumping an AGN wind has; the computational problem is a secondary one JM: Done..}
Given the lack of knowledge about the actual type of clumping, we have adopted
a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep[e.g.][]{hamann1998,hilliermiller1999,hamann2008}. 
The underlying assumption of microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that, {\em for fixed temperature},
 processes that are linear in density, such as electron scattering, are unchanged, 
as $f_V$ and $D$ will cancel out. However, any quantity that scales with the square of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.
In our models, the temperature is not fixed, and is instead set by balancing heating and 
cooling in a given cell. In the presence of an X-ray source, this thermal balance is 
generally dominated by bound-free heating and line cooling. The main effect of including 
clumping in our modelling is that it moderates the ionization state due to the increased 
density. This allows an increase in the ionizing luminosity, amplifying the amount of
bound-free heating and also increasing the competing line cooling term
(thermal line emission).

This treatment is necessarily simple; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, this parameterization allows simple estimates
of the effect clumping might have on the ionization state and emergent 
line emission.

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an eddington fraction of $0.2$; see section~\ref{sec:xray}).

We then evaluated these models based on whether their synthetic spectra exhibited the 
following properties of quasars and BALQSOs:

\begin{itemize}
\item UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)
\item Line emission emerging at low inclinations, with $EW\sim40\AA$ in \civ\ \citep[e.g. ][]{shen2011}
\item H recombination lines with $EW\sim50\AA$ in \la\ \citep[e.g. ][]{shen2011}
\item  \mg\ and \al\ (LoBAL) absorption features with $BI > 0$ at a small subset of BAL angles
\item Verisimilitude with quasar composite spectra.
\end{itemize}
Where $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
\begin{equation}
BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
\end{equation}
The constant $C=0$ everywhere, unless the normalized flux
has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
whereby $C$ is set to $1$.

@@

In the next section, we present one of the most promising models,
which we refer to as the fiducial model, and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
We then discuss the sensitivity to key parameters in section 5.
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

\begin{table}
\begin{tabular}{p{2cm}p{1cm}p{1cm}p{1cm}p{1cm}}
Parameter & \multicolumn{4}{|l|}{Grid Point Values}  \\
\hline \hline 
$r_{min}$ 	&	 $6r_{g}$ & $60r_{g}$ & \multicolumn{2}{|l|}{$300r_{g}$} \\ 
$\theta_{min}$ 	&	 $40^{\circ}$ & $55^{\circ}$ & \multicolumn{2}{|l|}{$70^{\circ}$} \\ 
$R_v$  	        &	 $10^{18}$cm & \multicolumn{3}{|l|}{$10^{19}$cm} \\ 
$\alpha$ 	&	 $0.5$ & $0.6$ & $0.75$ & $1.5$ \\
$f_V$ 	&	 $0.01$ & $0.1$ & \multicolumn{2}{|l|}{$1$}  \\
\hline 
\end{tabular}
\caption{The grid points used in the parameter search.}
\label{grid_table}
\end{table}

@@

\section{Results From A Fiducial Model}
Here we describe the results from a fiducial model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN. 
% We also examine trends with inclination in the synthetic spectra, 
% both in terms of the range of ionization states of the absorption lines and equivalent widths of the emission lines.

>>>>

\title
[Testing Quasar Unification]{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Testing Quasar Unification: Radiative Transfer in Clumpy Winds
%Unifying quasars with clumpy wind models
}

@@

Monte Carlo radiative transfer and photoionization code PYTHON.  
Previously, we have shown that we could produce synthetic spectra 
resembling those of observed broad absorption line (BAL) quasars 
for plausible wind models, but only if the X-ray luminosity was limited to about  
10$^{43}$~erg~s$^{-1}$. Here, we investigate the  degree to which
clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame ultraviolet (UV) 
that have the characteristics of quasars with X-ray luminosities as 
large as 10$^{45}$ erg s$^{-1}$. 
We conclude, perhaps not surprisingly, that clumping does maintain the ionization 
state of the wind necessary for strong BAL features in the rest-frame UV at more 
realistic X-ray luminosities. We examine the X-ray properties of our fiducial clumped model

@@

excited line emission in, e.g., \la and \civ\, to emerge at the low inclination, 
`Type 1 quasar-like' angles. At the highest inclinations, the synthetic spectra 
possess prominent \mg\ and \al\ BALs, the absorption features seen in LoBAL quasars. 
Despite these successes, we are unable to reproduce all the emission lines seen in
quasar spectra with the correct equivalent width 
ratios in any of the models in our initial simulation grid,
and we find an angular dependence of emission line equivalent width despite the 
remarkable similarities in the observed spectral properties of BAL and non-BAL quasars.

@@

ionization states seen in AGN, but a full quantitative agreement 
with quasar and BALQSO properties is not forthcoming.
% the remarkably uniform emission line properties 
% seen in BAL and non-BAL quasar composites, as we find a marked increase
% in emission line equivalent widths at high inclination `BALQSO-like' angles. 
% Overall, our work suggests that BAL outflows can produce significant broad emission
% lines and that accretion disc winds naturally reproduce the range of
% ionization states seen in AGN.
% This suggests that geometric unification
% pose a specific challenge to models in which an equatorial outflow
% rises from a limb-darkened, foreshortened accretion disc.

@@

Due to their ubiquitous nature,
disc winds offer a natural way to {\em unify} much of the
diverse phenemonology of luminous AGN and quasars \citep[e.g.][]{MCGV95, elvis2000}. 
Depending on viewing angle, an observer 
may see a BALQSO or normal `Type 1' quasar.
Within this geometric unification framework, the broad-line region (BLR) can 
correspond either to the dense wind base or clumps embedded
in the outflow. 
% Indeed, \cite{elitzur2014} show that a disc-wind BLR scenario
% naturally explains the emission line evolution of AGN.
A biconical wind model can also readily explain the various sub-classifications of BALQSOs: 
HiBALQSOs, which only exhibit high ionization line absorption; LoBALQSOs, which also show
absorption in lower ionization state species such as \mg\ and \al; and
FeLoBALQSOs, which show further absorption in Fe~\textsc{ii} and \textsc{iii}.
In unified geometric models, this is generally attributed to ionization stratification
of the outflow \citep[e.g.][]{elvis2000}.

@@

We have been engaged in a project to determine whether it is possible to simulate the properties of the spectra of AGN, including BALQSOs, using simple kinematic prescriptions for biconical disc winds using a Monte Carlo radiative transfer (MCRT) code that calculates the ionization structure of the wind and simulates the spectra from such a system \citep[][hereafter H13 and H14]{simlong2008,sim2010,higginbottom2013,H14}.  The results have been encouraging in the sense that in H13, we showed we could produce simulated spectra that resembled that of BALQSOs, as long as the luminosity of the X-ray source was relatively low, of order \POW{43}{\LUM} and the mass loss rate was relatively high, of order the mass accretion rate.  However, at higher X-ray luminosities, the wind was so ionized that UV absorption lines were not produced.  In addition, and in part due to limitations in our radiative transfer code, the model failed to produce spectra with strong emission lines at any inclination angle.

@@

For this study, we use the MCRT code \py 
\footnote{Named {\sl c. 1995}, predating the inexorable rise of a certain progamming language.} we have developed to carry out our 
radiative transfer and photoionization simulations in non-local-thermodynamic-equilibrium 
(non-LTE). The code can be used to model a variety of
disc-wind systems; applications have included accreting white dwarfs 
(Long \& Knigge 2002, hereafter LK02; Noebauer et al. 2010; 
Matthews et al. 2015, hereafter M15), young-stellar objects 
\citep{simmacro2005} and quasars/AGN (H13, H14).\nocite{noebauer, M15, LK02}

The code operates as follows:   This outflow is discretized into $n_x \times n_z$ cells in a 2.5D
cylindrical geometry with azimuthal symmetry. From some initial conditions in each cell, the code first calculates the ionization structure of the wind in a series of iterations.  Each iteration in an ``ionization cycle'' consists of generating photons, actually photon packets, 
from an accretion disc and central object, and calculating how these photon 
bundles scatter through the wind (eventually escaping the outflow or hitting the disk). 
The ionization and temperature structure is updated based on the properties of the 
radiation field in each cell, and the process is repeated. The radiative transfer and thermal
balance is carried out according to the principle of radiative equilibrium, 
which assumes both statistical equilibrium and that radiative heating balances radiative cooling.
Once the ionization structure has converged, it is held fixed, 
and synthetic spectra are generated at specific inclination 
angles in a series of "spectral cycles". LK02 provide a more detailed  
description of the original code; various improvements have been made 
since then and are described by \cite{simmacro2005}, H13 and M15.  
%A full description of the code will be published elsewhere; 
We focus here on the specific changes made for this study
intended to improve the ionization calculation of H and He
and to allow for clumping in the wind.
%These studies contain extensive detail on the code,  so we only briefly describe the key elements of the global ionization calculation and other important aspects.

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}). 
which are reprocessed by `macro-atoms' with associated transition probabilities.
% according to radiative equilibrium, which assumes both statistical equilibrium 
% and that radiative heating balances radiative cooling.
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15. In this scheme, the energy packets interact with either two-level 
`simple atoms' or full macro-atoms. 
This allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple atoms (see M15). In this study,
only H and He are treated as a macro-atom, because 
we expect recombination to be important
in determining their level populations and resultant line emission, 
and because we are especially interested in the contribution to 
AGN spectra of \LA.  H13 treated all atoms in a two-level approximation.  
% {\bf KSL:
% I thought we were also doing He. JM: yep, corrected.}

@@

In order to calculate the photoionization rate, 
we model the SED in a grid cell using the technique described by H13. In this scheme,
the mean intensity, $J_{\nu}$ in a series of $n$ bands is modeled as either a power law or exponential
in frequency $\nu$, with the fit parameters deduced from band-limited radiation field estimators.
This allows the calculation of a photoionization rate estimator. Ion abundances are
then calculated by solving the rate equations between ions. We include collisional ionization and photoionization balanced with radiative, 
dielectronic and collisional (three-body) recombination.
As in M15, we use a dilute Boltzmann approximation to calculate 
the populations of levels for simple-ions. This dilute approximation 
is not required for macro-atom levels. 
% \begin{equation}
% J_{\nu,i}=K_{pl}\nu^{\alpha_{pl}},
% \end{equation}
% for a band $i$, or an exponential 
% \begin{equation}
% J_{\nu,i}=K_{exp} e^{(-h\nu/k T_{exp})}.
% \end{equation}
% Here, $K_{exp}$, $K_{pl}$, $T_{exp}$ and $\alpha_{pl}$ are spectral fit parameters
% deduced from the band-limited radiation field estimators.
% The ionization rate out of ion $j$ can then be written as 
% \begin{equation}
% R_{j,j+1}(J)= 
% \displaystyle{
% n_j \left(C_{j} n_e + 
% \sum_{band~i=0}^{n}~
% {\int_{\nu_i}^{~\nu_i+1} \frac{4 \pi J_{\nu,i}\sigma_j(\nu)}  {h \nu} d\nu}
% \right),}
% \end{equation}
% where $\sigma_j$ is the photoionization cross-section, $n_e$
% is the electron density and $C_{j}$
% represents the collisional ionization coefficient.
% The recombination rate {\em into} ion $j$ is given by 
% \begin{equation}
% R_{j+1,j}(T_e) = (\alpha^j_{RR} + \alpha^j_{CR}) n_{j+1} n_e,
% \end{equation}
% where each $\alpha^j$ here is the recombination rate coefficient into the ground state of ion $j$.
% The subscripts denote radiative and 
% collisional (three-body) recombination, respectively. 
% $T_e$ is the electron temperature.
% %% We neglect recombination to and from excited states in the simple-ion calculation. 
% For simple-ions, we use a dilute Boltzmann approximation to calculate 
% the population of level $k$ in ionic stage $j$,
% \begin{equation}
% \frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R).
% \end{equation}
% Here $z_j$ is the partition function of ionic stage $j$,
% $T_R$ is the effective radiation temperature,
% $E_k$ is the energy difference between level $k$ and the ground state,
% and $g_k$ is the statistical weight of level $k$. 
% We stress that this approximation is not required for ions
% treated as macro-atoms. 
% {\bf KSL: 
% It seems a bit odd that you explain what we do with simple ions our old approach, but don't say what is done for macro-atoms,.  You only say it is not the other.
% JM: We've shortened the section on simple-ions as agreed..}

@@

In our previous modelling efforts, we assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
A clumpy outflow offers a possible solution to the so-called `over-ionization problem' in 
quasar and AGN outflows \citep{hamann2013}. This is the main motivation for incorporating 
clumping into our model.

Implementing a treatment of clumping is challenging, for
two main reasons. First, the physical scale lengths and density contrasts 
in AGN outflows are not well-constrained from observations.  
Second, there are significant computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. 
% Second, the addition of multiple additional degrees of
% freedom in the model results in significantly wider parameter space.
% Unfortunately, the physical scale lengths and density contrasts 
% associated with these parameters are not well-constrained from observations.  
% {\bf KSL: Reverse the order of these difficulties.  The most important is that it is not clear what kind of clumping an AGN wind has; the computational problem is a secondary one JM: Done..}
Given the lack of knowledge about the actual type of clumping, we have adopted
a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep[e.g.][]{hamann1998,hilliermiller1999,hamann2008}. 
The underlying assumption of microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that, {\em for fixed temperature},
 processes that are linear in density, such as electron scattering, are unchanged, 
as $f_V$ and $D$ will cancel out. However, any quantity that scales with the square of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.

In our models, the temperature is not fixed, and is instead set by balancing heating and 
cooling in a given cell. In the presence of an X-ray source, this thermal balance is 
generally dominated by bound-free heating and line cooling. The main effect of including 
clumping in our modelling is that it moderates the ionization state due to the increased 
density. This allows an increase in the ionizing luminosity, amplifying the amount of
bound-free heating and also increasing the competing line cooling term
(thermal line emission).
Our clumping treatment is necessarily simple; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, this parameterization allows simple estimates
of the effect clumping has on the ionization state and emergent 
line emission.

@@

Using this prescription, we conducted a limited parameter
search over a 5-dimensional parameter space involving the 
variables $r_{min}$, $\theta_{min}$, $f_V$, $\alpha$ and $R_v$.
The grid points are shown in table 1.
The aim here was to first fix $M_{BH}$ and $\dot{m}$ to their H13 values,
and increase $L_X$ to $10^{45}$~erg~s$^{-1}$ (a more realistic value for a 
quasar of $10^9M_\odot$ and an Eddington fraction of $0.2$; see section~\ref{sec:xray}).

We then evaluated these models based on 
how closely their synthetic spectra reproduced the 
following properties of quasars and BALQSOs:

\begin{itemize}
\item UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)
\item Line emission emerging at low inclinations, with $EW\sim40$\AA in \civ\ \citep[e.g. ][]{shen2011}
\item H recombination lines with $EW\sim50$\AA in \la\ \citep[e.g. ][]{shen2011}
\item  \mg\ and \al\ (LoBAL) absorption features with $BI > 0$ at a subset of 
BAL viewing angles
\item Verisimilitude with quasar composite spectra.
\end{itemize}
Here $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
\begin{equation}
BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
\end{equation}
The constant $C=0$ everywhere, unless the normalized flux
has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
whereby $C$ is set to $1$.

@@

In the next section, we present one of the most promising models,
which we refer to as the fiducial model, and discuss
the various successes and failures with respect to the above criteria.
This allows us to gain insight into fundamental geometrical 
and physical constraints and assess the potential for unification. 
We then discuss the sensitivity to key parameters in section~\ref{sec:param_sens}.
The full grid, including output synthetic spectra and plots can be found at
\url{jhmatthews.github.io/quasar-wind-grid/}.

@@

\begin{table}
\begin{tabular}{p{2cm}p{1cm}p{1cm}p{1cm}p{1cm}}
Parameter & \multicolumn{4}{|l|}{Grid Point Values}  \\
\hline \hline 
$r_{min}$ 	&	 $60r_{g}$ & $180r_{g}$ & \multicolumn{2}{|l|}{$300r_{g}$} \\ 
$\theta_{min}$ 	& $55^{\circ}$ & \multicolumn{3}{|l|}{$70^{\circ}$} \\ 
$R_v$  	        &	 $10^{18}$cm & \multicolumn{3}{|l|}{$10^{19}$cm} \\ 
$\alpha$ 	&	 $0.5$ & $0.6$ & $0.75$ & $1.5$ \\
$f_V$ 	&	 $0.01$ & \multicolumn{3}{|l|}{$0.1$}  \\
\hline 
\end{tabular}
\caption{The grid points used in the parameter search.
The sensitivity to some of these parameters is discussed 
further in section~\ref{sec:param_sens}}
\label{grid_table}
\end{table}

@@

\section{Results and Discussion From A Fiducial Model}
Here we describe the results from a fiducial model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN. 
% We also examine trends with inclination in the synthetic spectra, 
% both in terms of the range of ionization states of the absorption lines and equivalent widths of the emission lines.

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Fiducial Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{m}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{erg~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{m}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
%%$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f_V$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters 
used in the fiducial model, as defined in the text and figure 1.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk.}
\label{wind_param}
\end{table}
